---
title: "Replication Study"
author: "Sarah Jensen"
date: "11/11/2021"
output:
  word_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
In this project, I attempt to replicate the results of a particularly influential paper on goal setting and unethical behavior (Schweitzer, Ordonez, & Douma, 2004).

First, I'll load libraries and the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load libraries
library (dplyr)
library (ggplot2)
library (tidyverse)
library (psych)
library(readr)
library (ggpubr)
library(readxl)
library(lsmeans)
library (car)
library (rstatix)
library (rcompanion)
library (stringr)

#load data
## I removed Prolific ID's, IP addresses before importing here
## I removed a row of headers that wasn't needed
## I removed "test"/preview submissions

df <- read_excel("~/GitHub/Replication_Study/repdata_12.9.21.xlsx")
View(df)

```

Clean the data

```{r, include=FALSE}
## select all the columns I need for analysis (mostly take out timing questions)
df <- df %>%
  filter (comp1 == "2m3v") %>%
  select (PID, comp1, comp2, practice1, practice2, r1, r2, r3, r4, r5, r6, r7, goalcommit_1, goalcommit_2, goalcommit_3, goalcommit_4, goalcommit_5, goalcommit_6, goalcommit_7, goalcommit_8, report_1, report_2, report_3, report_4, report_5, report_6, report_7, bonus_report_reward, goal_report_mere, report_dyb, age, gender, race, work_exp, sanitycheck, anythingodd, condition)
  

##recode goal commitment scores (items 1,2,3,4 & 6 are reverse coded)
df$goalcommit_1<-dplyr::recode(df$goalcommit_1, 'Strongly Disagree'=5, 'Moderately Disagree'=4, 'Neutral'=3, 'Moderately Agree'=2,'Strongly Agree'=1)

df$goalcommit_2<-dplyr::recode(df$goalcommit_2, 'Strongly Disagree'=5, 'Moderately Disagree'=4, 'Neutral'=3, 'Moderately Agree'=2,'Strongly Agree'=1)

df$goalcommit_3<-dplyr::recode(df$goalcommit_3, 'Strongly Disagree'=5, 'Moderately Disagree'=4, 'Neutral'=3, 'Moderately Agree'=2,'Strongly Agree'=1)

df$goalcommit_4<-dplyr::recode(df$goalcommit_4, 'Strongly Disagree'=5, 'Moderately Disagree'=4, 'Neutral'=3, 'Moderately Agree'=2,'Strongly Agree'=1)

df$goalcommit_5<-dplyr::recode(df$goalcommit_5, 'Strongly Disagree'=1, 'Moderately Disagree'=2, 'Neutral'=3, 'Moderately Agree'=4,'Strongly Agree'=5)

df$goalcommit_6<-dplyr::recode(df$goalcommit_6, 'Strongly Disagree'=5, 'Moderately Disagree'=4, 'Neutral'=3, 'Moderately Agree'=2,'Strongly Agree'=1)

df$goalcommit_7<-dplyr::recode(df$goalcommit_7, 'Strongly Disagree'=1, 'Moderately Disagree'=2, 'Neutral'=3, 'Moderately Agree'=4,'Strongly Agree'=5)

df$goalcommit_8<-dplyr::recode(df$goalcommit_8, 'Strongly Disagree'=1, 'Moderately Disagree'=2, 'Neutral'=3, 'Moderately Agree'=4,'Strongly Agree'=5)
  
##change things to numeric
df$report_1 <- as.numeric(df$report_1)
df$report_2 <- as.numeric(df$report_2)
df$report_3 <- as.numeric(df$report_3)
df$report_4 <- as.numeric(df$report_4)
df$report_5 <- as.numeric(df$report_5)
df$report_6 <- as.numeric(df$report_6)
df$report_7 <- as.numeric(df$report_7)

```

Data manipulation:
-Create dictionaries of the eligible words and create "actual" columns
-Calculate totals for reported and actual words
-Calculate difference scores

```{r, include=FALSE}
#Make dictionaries of all words for each letter string

#r1: AEDKBUG
words1 <- c("debug","badge", "baked", "budge", "daube", "egad", "baud", "gaud", "bead", "beak", "beau", "daub", "duke", "bake", "abed", "kagu", "aged", "ague", "gude", "bade", "dub", "due", "dug", "keg", "kea", "gab", "gad", "ged", "deb", "bad", "bag", "age", "bed", "beg", "bud", "bug", "dab", "dag", "dak", "auk", "ad", "ba", "ab", "be")
##words removed: bedu, gaed, kae, kue, eau, kab (alternative for cab), uke (abbrv. for ukelele), da

#r2: OASFKEV
words2 <- c("fakes", "faves", "fovea", "soave", "aves", "avos", "fake", "fave", "foes", "kafs", "keas", "kefs", "koas", "oafs", "oaks", "okas", "okes", "safe", "sake", "save", "soak", "sofa", "soke", "vase", "ask", "ave", "avo", "foe", "kaf", "kas", "kea", "kef", "koa", "kos", "oaf", "oak", "oka", "oke", "ova", "sea", "ska", "vas", "voe", "ae", "as", "fa", "ka", "of", "os", "so")
##words removed: foveas, oaves, kaes, kvas, voes, efs, fas, fes, kae, kev, oes, ose, sae, ef, es, fe, oe
##kept fave and faves as slang...

#r3: OELBJAM
words3 <- c("jambe", "amble", "amole", "blame", "ebola", "able", "alme", "aloe", "ambo", "bale", "balm", "beam", "bema", "blae", "bola", "bole", "jamb", "lamb", "lame", "loam", "lobe", "male", "meal", "mola", "mole", "olea", "alb", "ale", "bal", "bam", "bel", "boa", "elm", "jab", "jam", "job", "joe", "lab", "lam", "lea", "lob", "mae", "mel", "moa", "mob", "oba", "ab", "ae", "al", "am", "ba", "be", "el", "em", "jo", "la", "lo", "ma", "me", "mo")
##words removed: blam, jole, mabe, mol, obe, ole, bo, oe, om

#r4: UADQWER
words4 <- c("dewar", "quare", "wader", "awed", "dare", "dear", "draw", "drew", "dura", "dure", "quad", "read", "rude", "rued", "urea", "wade", "ward", "ware", "waur", "wear", "are", "awe", "daw", "dew", "due", "ear", "era", "qua", "rad", "raw", "red", "rue", "urd", "wad", "war", "wed", "wud", "ad", "ae", "aw", "ed", "er", "re", "we")
## words removed: wared, waur (Scottish), eau, wae (Scottish), ar, da, de (spanish/french)

#r5: EASCKIY
words5 <- c("cakes", "cakey", "saice", "skiey", "yacks", "yikes", "aces", "cask", "sake", "asci", "cays", "saki", "yack", "ayes", "easy", "keas", "sice", "yaks", "cake", "eyas", "keys", "sick", "yeas", "ices", "sika", "case", "icky", "sack", "syce", "ace", "ice", "key", "sec", "yes", "ick", "ask", "icy", "sic", "aye", "sac", "ska", "kas", "ski", "cay", "say", "sky", "cis", "kea", "sea", "yak", "yea", "ae", "ai", "as", "ay", "is", "ka", "si", "ya", "ye")
##words removed: casky, kae, kaes, syke, kays, caky, kyes, kye, sike, ais, kis, sei, ays, sae, kay, es, ki

#r6: OELHMAZ
words6 <- c("almeh", "amole", "azole", "haole", "hazel", "hemal", "mahoe", "mohel", "ahem", "alme", "aloe", "haem", "hale", "halm", "halo", "hame", "haze", "heal", "helm", "helo", "hole", "holm", "home", "lame", "laze", "loam", "male", "maze", "meal", "mola", "mole", "olea", "zeal", "zoea", "ale", "lam", "azo", "lea", "ohm", "elm", "mae", "zoa", "ham", "mel", "hao", "mho", "hem", "moa", "ae", "ah", "al", "am", "eh", "el", "em", "ha", "he", "hm", "la", "lo", "ma", "me", "mo", "oh", "om")
##words removed: zoeal, mol, lez (offensive term for lesbian), ole, hae, oe, za

#r7: AODJGIP
words7 <- c("podia", "agio", "dopa", "goad", "jiao", "paid", "ado", "ago", "aid", "apo", "dag", "dap", "dig", "dip", "dog", "gad", "gap", "goa", "god", "jag", "jig", "jog", "oda", "pad", "pod", "poi", "ad", "ag", "ai", "do", "gi", "go", "id", "jo", "od", "pa", "pi")
## words removed: gadjo, pagod, apod, gadi, padi, gip (offensive), pia, da, di, oi, op (used as op art)

#make new columns for the actual number of words created in each round using the dictionaries above
#the final df is new7
new1 <- within(
  df,
  actual_1 <- sapply(
    gregexpr(paste0(words1, collapse = "|"), r1, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new2 <- within (
  new1,
  actual_2 <- sapply(
    gregexpr(paste0(words2, collapse = "|"), r2, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new3 <- within (
  new2,
  actual_3 <- sapply(
    gregexpr(paste0(words3, collapse = "|"), r3, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new4 <- within (
  new3,
  actual_4 <- sapply(
    gregexpr(paste0(words4, collapse = "|"), r4, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new5 <- within (
  new4,
  actual_5 <- sapply(
    gregexpr(paste0(words5, collapse = "|"), r5, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new6 <- within (
  new5,
  actual_6 <- sapply(
    gregexpr(paste0(words6, collapse = "|"), r6, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)
new7 <- within (
  new6,
  actual_7 <- sapply(
    gregexpr(paste0(words7, collapse = "|"), r7, ignore.case = TRUE),
    function(x) sum(x > 0)
    )
)

##create totals for reported and actual words
##create difference score between actual and reported for each of the 7 rounds and the total
data1 <- new7 %>%
  mutate (reporttotal=rowSums(select(.,report_1:report_7)), actualtotal=rowSums(select(., actual_1:actual_7)), differencetotal=actualtotal- reporttotal, diff_1=report_1-actual_1, diff_2=report_2-actual_2, diff_3=report_3-actual_3, diff_4=report_4-actual_4, diff_5=report_5-actual_5, diff_6=report_6-actual_6, diff_7=report_7-actual_7) 

#make new columns for the total number of words overreported and underreported
##differencetotal_pos and differencetotal_neg will sum the words
##COUNT_POS and COUNT_NEG will count the number of rounds in which words were overcounted
data2 <- data1 %>%
  select(., diff_1:diff_7) %>%
  mutate(differencetotal_pos = rowSums(mutate_all(., ~if_else(.x > 0, .x, NA_real_)), na.rm = TRUE),
         COUNT_POS = rowSums(. > 0, na.rm = TRUE)) %>%
  mutate(differencetotal_neg = rowSums(mutate_all(., ~if_else(.x < 0, .x, NA_real_)), na.rm = TRUE),
         COUNT_NEG = rowSums(. < 0, na.rm = TRUE)) %>%
  select (., diff_7:COUNT_NEG)

#combine the datasets and drop "diff&7" duplicate column
data3 <- cbind (data1, data2) %>%
  subset(., select=which(!duplicated(names(.)))) 

#overwrite data1
data1 <- data3

```

Explore the data + demographics

```{r}
#demographics
data1$age <- as.numeric(data1$age)
mean(data1$age, na.rm = TRUE) #36.77
sd(data1$age, na.rm = TRUE) #13.43
table (data1$gender) #51% female (86/168)
table (data1$race) #76% white (128/168)
data1$work_exp <- as.numeric(data1$work_exp)
mean(data1$work_exp, na.rm = TRUE) #14.92


#summarize the totals by condition
summary1 <- data1 %>%
  group_by (condition) %>%
  get_summary_stats (reporttotal, actualtotal, differencetotal, type="mean_sd")

print(summary1)

#compare words reported in each group 
oneway.test(reporttotal ~ condition, data = data1, var.equal = TRUE)
oneway_result <- aov(reporttotal ~ condition, data = data1)
TukeyHSD(oneway_result)

#compare number of words actually generated in each group
oneway.test(actualtotal ~ condition, data = data1, var.equal = TRUE)
oneway_result <- aov(actualtotal ~ condition, data = data1)
TukeyHSD(oneway_result)

```
### Analysis for H1: People with specific, unmet goals will be more likely to overreport their performance compared with those who don't have specific goals.
```{r}
#make new df for comparing number of words in the goal conditions to the DYB condition 
ttest_df <- data1 %>%
  select (condition, reporttotal, actualtotal, differencetotal_pos, differencetotal_neg) %>%
  mutate (condition =recode_factor(condition, doyourbest = "doyourbest", meregoal= "goalavg", rewardgoal= "goalavg"))

#t-test comparing actual number of words
t.test(actualtotal ~ condition, ttest_df, var.equal = FALSE)

#t-test comparing reported number of words
t.test(reporttotal ~ condition, ttest_df, var.equal = FALSE)

#t-test comparing difference scores for overreport and underreport
t.test(differencetotal_pos ~ condition, ttest_df, var.equal=FALSE)
t.test(differencetotal_neg ~ condition, ttest_df, var.equal=FALSE)

#summary table of stats
summary_ttest <- ttest_df %>%
  group_by (condition) %>%
  get_summary_stats (reporttotal, actualtotal, differencetotal_neg, differencetotal_pos, type="mean_sd")

print(summary_ttest)

#percentages of people who overstated at least once
overstated <- data1 %>%
  select (condition, differencetotal_pos) %>%
  group_by(condition) %>%
  summarize (total=sum(differencetotal_pos>=1))
##DYB 36.2%
##Meregoal 50%
##Reward 53.7%


#I ran the chi-squared to see if there are differences beteween overreports under H3,  but it should be here instead. The result is under (chisq3)

#Also tried to do this for the 2 levels (goal conditions and DYB)
anotherdf <- data1 %>%
  select (condition, reporttotal, actualtotal, differencetotal_pos, differencetotal_neg, actual_1:actual_7, diff_1:diff_7) %>%
  mutate (condition =recode_factor(condition, doyourbest = "doyourbest", meregoal= "goalavg", rewardgoal= "goalavg"))
  
actual_cols <- 6:12
diff_cols <- 13:19
anotherdf$didtheyoverreport <- rowSums(anotherdf[diff_cols] * as.numeric(anotherdf[diff_cols] != 0))

anotherdf <- anotherdf %>%
  filter (didtheyoverreport >=0) %>%
  mutate (didtheyoverreport2 = ifelse(didtheyoverreport == 0, 0, 1))

table5 <- table(anotherdf$condition, anotherdf$didtheyoverreport2)
print(table5)

chisq4 <- chisq.test(anotherdf$condition, anotherdf$didtheyoverreport2, correct=FALSE)
(chisq4)

#run one-way ANOVA  
oneway.test (differencetotal_pos ~ condition, data = chisq_df1, var.equal = TRUE)
oneway_result <- aov(differencetotal_pos ~ condition, data = chisq_df1)
TukeyHSD(oneway_result)

```
### Analysis for H2: People with unmet reward goals will be more likely to overstate their performance than people with unmet mere goals.
```{r}
#t-test that examines difference score in rewardgoal vs meregoal conditions
ttest_df2 <- data1 %>%
  filter (condition != "doyourbest") %>%
  select (condition, reporttotal, differencetotal_pos)

t.test(differencetotal_pos ~ condition, ttest_df2, var.equal = FALSE)

#make table to look at proportions of people - HAVE NOT FINISHED THIS


#chi-squared to examine over reports in just the goal conditions
chisq_df2 <- data1 %>%
  filter (condition != "doyourbest") %>%
  select (condition,differencetotal_pos, differencetotal_neg) %>%
  group_by (condition) %>%
  summarize (overreport=sum(differencetotal_pos >= 1),
             underreport=sum(differencetotal_neg <=-1)) %>%
  gather (overreport, underreport, key=report, value=words)

table3 <- data.frame(N= c(56, 54),
                     condition= c("meregoal", "rewardgoal"),
                     report= c("overreport", "underreport"),
                     value1= c(28, 23),
                     value2= c(29, 27))
table3 <- table3 %>%
  mutate(overreport_prop=overreport/N*100,
         underreport_prop=underreport/N*100) 

#OLD
table1 <- left_join (chisq_df2, table1, copy=TRUE, by="condition")
table2 <- table1 %>%
  mutate(overreport_prop=overreport/N*100,
         underreport_prop=underreport/N*100) %>%
  subset(select=c("condition","overreport_prop","underreport_prop"))

table(chisq_df2$overreport, chisq_df2$underreport)
conTable(table2)

chisq2 <- chisq.test(chisq_df2$condition, chisq_df2$report, correrct=FALSE)
(chisq2)
chisq2 <- chisq.test(table(chisq_df2))

```
### Analysis H3: People who fail to reach their goals by a small margin will be more likely to falsely claim to have reached their goals than people who fail to reach their goals by a large margin.
```{r}
#make df keeping all rounds & totals for only goal conditions, filter out people who made 9 or more words (so less than or equal to 8 words)
#for the remaining data, select the actual columns to find those who overstated once and those who accurately stated once
#make report_correct & report_incorrect variables

df3 <- data1 %>%
  select (diff_1:diff_7, condition, differencetotal_pos, actual_1:actual_7) %>%
  filter (condition != "doyourbest") %>%
  filter (actual_1<=8) %>%
  filter (actual_2<=8) %>%
  filter (actual_3<=8) %>%
  filter (actual_4<=8) %>%
  filter (actual_5<=8) %>%
  filter (actual_6<=8) %>%
  filter (actual_7<=8) %>%
  filter (differencetotal_pos !=0) %>%
  filter (diff_1 ==0 | diff_2==0 | diff_3==0 | diff_4==0 | diff_5==0| diff_6==0 | diff_7==0) 

#make new column for the number of words reported in the correctly reported rounds
#make new column for the number of words reported in the incorrectly reported rounds
actual_cols <- 10:16
diff_cols <- 1:7
df3$report_correct <- rowSums(df3[actual_cols] * as.numeric(df3[diff_cols] == 0))
df3$report_incorrect <- rowSums(df3[actual_cols] * as.numeric(df3[diff_cols] != 0))

#run paired t-tests
t.test(df3$report_correct, df3$report_incorrect, paired = TRUE, alternative = "two.sided")

table_df3 <- df3 %>%
  get_summary_stats (report_correct, report_incorrect, type= "mean_sd")

#for the cases in which 8 words were created, 
df4 <- data1 %>%
  select (diff_1:diff_7, condition, differencetotal_pos, actual_1:actual_7) %>%
  filter (actual_1==8 |actual_2==8 |actual_3==8 |actual_4==8 |actual_5==8 |actual_6==8 |actual_7==8)

##make a new column- 0=didn't overreport, 1=overreported at least once
actual_cols <- 10:16
diff_cols <- 1:7
df4$didtheyoverreport <- rowSums(df4[diff_cols] * as.numeric(df4[actual_cols] == 8))

##filter out underreports & make a new column coding overreports as 1 and not overeports as 0. Then make a contingency table
df4 <- df4 %>%
  filter (didtheyoverreport >=0) %>%
  mutate (didtheyoverreport2 = ifelse(didtheyoverreport == 0, 0, 1))

table4 <- table (df4$condition, df4$didtheyoverreport2)

chisq3 <- chisq.test(df4$condition, df4$didtheyoverreport2, correct=FALSE)
(chisq3)
  
```

Goal commitment scores analysis

```{r, echo=FALSE}
#make a new df and run ordinary least squares regression
df5 <- data1 %>%
  filter (condition != "doyourbest") %>%
  select (condition, actualtotal, goalcommit_1:goalcommit_8) %>%
  mutate (gc_score = rowSums(select(., goalcommit_1:goalcommit_8)))

ols_test <- lm(actualtotal ~ gc_score * condition, data = df5) 
summary(ols_test)

#now run it with just questions 4-8 (short goal commitment scale)
df6 <- data1 %>%
  filter (condition != "doyourbest") %>%
  select (condition, actualtotal, goalcommit_4:goalcommit_8) %>%
  mutate (gc_score1 = rowSums(select(., goalcommit_4:goalcommit_8)))

short_ols_test <- lm(actualtotal ~ gc_score1 * condition, data = df6) 
summary(short_ols_test)

```

Some data visualizations

```{r, echo=FALSE}
# barplot showing the average number of reported words per condition. 
##there is a higher average of reported words in the reward goal condition.
plot1 <- data1 %>%
  group_by (condition) %>%
  get_summary_stats (reporttotal, type="mean_sd") %>%
  ggplot (aes(condition, mean, fill=condition)) +
  geom_bar (stat = "identity") +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge()) +
  theme_pubr (base_size=12, legend="none") +
  scale_fill_grey() +
  labs (title="Number of Words Reported By Condition", y="Number of words", x="Condition")

(plot1)

#overstatement and overstatement mean scores by condition
dat3 <- subset (data1, differencetotal >=0) %>%
  group_by (condition) %>%
  summarize (diffT=mean(differencetotal))

dat4 <- subset (data1, differencetotal <=0) %>%
  group_by (condition) %>%
  summarize (diffT=mean(differencetotal))

p1 <- dat3 %>%
  ggplot(aes(x=condition, y=diffT, group=1))+
    geom_line ()+
    geom_point(size=2) +
    theme_pubr (base_size=12, legend="none") +
  labs (title="Average Number of Overreported and UnderReported Words by Condition", y="Words", x= " ")

p2 <- dat4 %>%
  ggplot(aes(x=condition, y=diffT, group=1))+
    geom_line ()+
    geom_point(size=2) +
    theme_pubr (base_size=12, legend="none") +
    labs (y="Words", x="Condition")

plot2 <- ggarrange(p1, p2, ncol = 1)

(plot2)

#Histogram showing the number of words overreported
plot3 <- ggplot (data1, aes(differencetotal_pos)) +
  geom_histogram(binwidth=1) +
  theme_pubr (base_size=12, legend="none") +
  scale_fill_grey() +
  labs (title="Overreports", y="Number of People", x="Number of overreports")
(plot3)

#Histogram showing the number of words underreported
plot4 <- ggplot (data1, aes(differencetotal_neg)) +
  geom_histogram(binwidth=1) +
  theme_pubr (base_size=12, legend="none") +
  scale_fill_grey() +
  labs (title="Underreports", y="Number of People", x="Number of underreports")
(plot4)

#Histogram showing the number of rounds in which people overreported
plot5<- ggplot (data1, aes(COUNT_POS)) +
  geom_histogram(binwidth=1) +
  theme_pubr (base_size=12, legend="none") +
  scale_fill_grey() +
  labs (title="Number of Rounds in which participants overreported", y="People who overreported", x="Rounds")
(plot5)

#Histogram showing the number of rounds in which people underreported
plot6 <- ggplot (data1, aes(COUNT_NEG)) +
  geom_histogram(binwidth=1) +
  theme_pubr (base_size=12, legend="none") +
  scale_fill_grey() +
  labs (title="Number of Rounds in which participants underreported", y="People who underreported", x="Rounds")
(plot6)
    
#table replicated from pg. 427
table427 <- data1 %>%
  select (condition, actual_1:actual_7) %>% 
  group_by (condition) %>% 
  get_summary_stats (actual_1, actual_2, actual_3, actual_4, actual_5, actual_6, actual_7, type="mean_sd")

print(table427)
          
```
